---
title: "More Mapping Presentation"
author: "Jenna Moscaritolo & Jung Hwa Yeom"
date: "11/11/2020"
output:
  revealjs::revealjs_presentation:
    self_contained: false
    reveal_options:
      slideNumber: true
      previewLinks: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Introduction

'hurricaneexposuredata' package

- Rain data
- Wind data
- Latitude & longitude


## Goal

To compare amount of rainfall of three consecutive hurricanes and their storm tracks.


## Plan

Because we wanted three hurricanes, we decided to take one for each year in 2009-2011.

- Step 1: Choosing the three hurricanes
- Step 2: Cleaning the data
- Step 3: Connecting the coordinates with the other data
- Step 4: Mapping the data


## Filtering Target Rain Data

After doing the initial extraction of the data, we filtered the data to have the rain data link with our desired hurricanes.

```{r, echo = FALSE}
pacman::p_load("tidyverse","drat","maps","tmap","sf","viridis","sp","hurricaneexposuredata",
               "hurricaneexposure","tmaptools","revealjs")
addRepo("geanders")
data("hurr_tracks") 
data("rain")

# Pulling map data
ob <- st_as_sf(map('county',plot=FALSE,fill=TRUE))
colnames(county.fips)[2] = 'ID'
ob <- merge(ob, county.fips, by="ID")

# Filtering target hurr_tracks data
dt_o <- hurr_tracks %>% filter(storm_id == "One-2009")
dt_e <- hurr_tracks %>% filter(storm_id == "Earl-2010")
dt_i <- hurr_tracks %>% filter(storm_id == "Irene-2011")

# Filtering target rain data
ra_o <- rain %>% filter(storm_id == "One-2009") %>% group_by(fips) %>%
  summarise('storm_id' = storm_id[1], 'precip' = sum(precip))
ra_e <- rain %>% filter(storm_id == "Earl-2010") %>% group_by(fips) %>%
  summarise('storm_id' = storm_id[1], 'precip' = sum(precip))
ra_i <- rain %>% filter(storm_id == "Irene-2011") %>% group_by(fips) %>%
  summarise('storm_id'=storm_id[1],'precip'=sum(precip))
```


## Using Coordinates for Hurricane Curve

We combined the data to produce a curve for the hurricane tracks of the three hurricanes using the coordinate data.

```{r, echo = FALSE}
# Changing fips to numerics
ra_o$fips <- as.numeric(ra_o$fips)
ra_e$fips <- as.numeric(ra_e$fips)
ra_i$fips <- as.numeric(ra_i$fips)

# Merging data by fips variable
ra_o <- merge(ob, ra_o, by="fips")
ra_e <- merge(ob, ra_e, by="fips")
ra_i <- merge(ob, ra_i, by="fips")

# Pulling the coordinates for each hurricane path
One <- cbind(dt_o$longitude,dt_o$latitude) %>%
  Line() %>% Lines(ID = 'One-2009') %>%
  list() %>% SpatialLines()
Earl <- cbind(dt_e$longitude,dt_e$latitude)%>%
  Line()%>%Lines(ID='Earl-2010')%>%
  list()%>%SpatialLines()
Irene <- cbind(dt_i$longitude, dt_i$latitude) %>%
  Line() %>% Lines(ID = 'Irene-2011') %>%
  list() %>% SpatialLines()
```


## Plotting the Maps

We used the tmap method because we were able to tailor the maps exactly how we wanted it to look.

```{r, echo = FALSE}
# Organize map data
ra_o <- rain %>% filter(storm_id == "One-2009") %>% group_by(fips) %>%
  summarise(storm_id=storm_id[1],precip=sum(precip))
ra.new.o <- as.data.frame(ra_o)

ra.new.o$fips <- as.numeric(ra.new.o$fips)
ra_e <- rain %>% filter(storm_id == "Earl-2010") %>% group_by(fips) %>%
  summarise(storm_id=storm_id[1],precip=sum(precip))
ra.new.e <- as.data.frame(ra_e)
ra.new.e$fips <- as.numeric(ra.new.e$fips)

ra_i <- rain %>% filter(storm_id == "Irene-2011") %>% group_by(fips) %>%
  summarise(storm_id=storm_id[1],precip=sum(precip))
ra.new.i <- as.data.frame(ra_i)
ra.new.i$fips <- as.numeric(ra.new.i$fips)

# Merging data by fips variable
ra.new.o <- merge(ob, ra.new.o, by="fips")
ra.new.e <- merge(ob, ra.new.e, by="fips")
ra.new.i <- merge(ob, ra.new.i, by="fips")
```

```{r}
# Plotting all three hurricane tracks and amount of rainfall
#tmap for Hurricane One
t_F_o = tm_shape(ra.new.o)+
  tm_polygons(col='precip',title="Rainfall (mm)")+
  tm_legend(position=c("right","bottom"))+
  tm_shape(One)+
  tm_lines(col='red')+
  tm_layout(main.title='One-2009',main.title.position = "center") 


#tmap for Hurricane Earl
t_F_e = tm_shape(ra.new.e)+
  tm_polygons(col='precip',title="Rainfall (mm)")+
  tm_legend(position=c("right","bottom"))+
  tm_shape(Earl)+
  tm_lines(col='red')+
  tm_layout(main.title='Earl-2010',main.title.position = "center")


#tmap for Hurricane Irene
t_F_i = tm_shape(ra.new.i)+
  tm_polygons(col='precip',title="Rainfall (mm)")+
  tm_legend(position=c("right","bottom"))+
  tm_shape(Irene)+
  tm_lines(col='red')+
  tm_layout(main.title='Irene-2011',main.title.position = "center") 
```


## Hurricane One Map

```{r, echo = FALSE}
t_F_o
```


## Hurricane Earl Map

```{r, echo = FALSE}
t_F_e
```


## Hurricane Irene Map

```{r, echo = FALSE}
t_F_i
```


## Problems

**Problem:** FEMA data was provided to try to match hurricanes with the 'hurricaneexposuredata' package but there were many storms that were not in one dataset that were in another.

**Solution:** We disregarded the FEMA data and found hurricanes that fit our description of being in three consecutive years.


## Conclusion

Hurricane One

- Medium precipitation level
- Low precipication levels over a wide space

Hurricane Earl

- Lower precipitation levels
- Similar track as Hurricane One

Hurricane Irene

- Higher precipitation levels
- Direct contact with land


## Thank you!

Questions?
